---
title: "Assignment_1"
author: "Melissa Jaffe"
date: "9/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
path <- "C:/Users/Melissa/Documents/Ecostats"
setwd(path)
knitr::opts_knit$set(root.dir = path)
```

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(scales)
library(tidyverse)
```

```{r}
clim_tree_data <- read.csv("tahoe_snotel_data.csv", header = T)
na.omit(clim_tree_data)
head(clim_tree_data)
sapply(clim_tree_data,class)
summary(clim_tree_data)
str(clim_tree_data)
```
#Question 1: Summarize the data by monthly normals (means) for each site
```{r}
subset_clim <- clim_tree_data %>% 
  group_by(site, month, year) %>%
  summarize(mean(tavg, na.rm = TRUE))
subset_clim

subset_clim_2 <- clim_tree_data %>% 
  group_by(site, elev, month, year) %>%
  summarize(mean(tavg, na.rm = TRUE))
subset_clim_2

head(subset_clim)

names(subset_clim)[names(subset_clim) == "mean(tavg, na.rm = TRUE)"] <- "t_average"
names(subset_clim_2)[names(subset_clim_2) == "mean(tavg, na.rm = TRUE)"] <- "t_average"

is.data.frame(subset_clim)
  
```
#Question 2: Produce univariate summary statistics for data. Produce univariate plots characterizing the data
```{r}
hist_clim <- ggplot(subset_clim, aes(t_average)) + geom_histogram() + facet_wrap(~site)
hist_clim

year_hist <- ggplot(subset_clim, aes(year)) + geom_histogram() + facet_wrap(~site)
year_hist #not all sites have all years

elev_dot <- ggplot(subset_clim_2, aes(x = site, y = elev)) + geom_point() 
elev_dot
```
#Question 3: Fit a linear model that relates the average temperature in May to elevation using data from the 16 sites. Do the same for average temperatures in December. What are the lapse rates from these months? What do these values mean? 
```{r}
new_data_1 <- clim_tree_data %>%
  group_by(site, month) %>%
  summarize(mean(elev, na.rm = TRUE), mean(tavg, na.rm = TRUE))
names(new_data_1)[names(new_data_1) == "mean(tavg, na.rm = TRUE)"] <- "t_average"
names(new_data_1)[names(new_data_1) == "mean(elev, na.rm = TRUE)"] <- "elev"
head(new_data_1)

may_data <- clim_tree_data %>%
 subset(clim_tree_data$month == "May") %>%
  group_by(site, elev) %>%
  summarize(mean(tavg, na.rm = TRUE))
names(may_data)[names(may_data) == "mean(tavg, na.rm = TRUE)"] <- "t_average"

dec_data <- clim_tree_data %>%
 subset(clim_tree_data$month == "December") %>%
  group_by(site, elev) %>%
  summarize(mean(tavg, na.rm = TRUE))
names(dec_data)[names(dec_data) == "mean(tavg, na.rm = TRUE)"] <- "t_average"

  
may_lm <- lm(may_data$elev~may_data$t_average)
summary(may_lm)

dec_lm <- lm(dec_data$elev~dec_data$t_average)
summary(dec_lm)

```
#Question 4: Plot the relationship
```{r}

may_plot <- ggplot(may_data, aes(x = elev, y= t_average)) + geom_point() +geom_smooth(method = lm)
may_plot

dec_plot <-  ggplot(dec_data, aes(x = elev, y= t_average)) + geom_point() +geom_smooth(method = lm)
dec_plot
#without ggplot
plot(may_data$elev, may_data$t_average, xlab="Elevation", ylab ="Temp Average")
abline(lm(may_data$elev~may_data$t_average), col = "red")

plot(dec_data$elev, dec_data$t_average, xlab="Elevation", ylab ="Temp Average")
abline(lm(dec_data$elev~dec_data$t_average), col = "red")

```
#Question 5: Produce diagnostic  plots characterizing the fit of the linear model and its assumptions
```{r}
may_resids <- resid(may_lm)
may_y_hat <- fitted(may_lm)
plot(may_y_hat, may_resids)
qnorm(may_resids)
plot(may_lm)

dec_resids <- resid(dec_lm)
dec_y_hat <- fitted(dec_lm)
plot(dec_y_hat, dec_resids)
qnorm(dec_resids)
plot(dec_lm)
```

